---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(ggpsychro)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    fig.align = "center",
    out.width = "60%",
    dev = "png",
    dpi = 100
)
```

# ggpsychro

<!-- badges: start -->
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/hongyuanjia/ggpsychro?branch=master&svg=true)](https://ci.appveyor.com/project/hongyuanjia/ggpsychro)
[![Travis build status](https://travis-ci.com/hongyuanjia/ggpsychro.svg?branch=master)](https://travis-ci.com/hongyuanjia/ggpsychro)
[![Codecov test coverage](https://codecov.io/gh/hongyuanjia/ggpsychro/branch/master/graph/badge.svg)](https://codecov.io/gh/hongyuanjia/ggpsychro?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/ggpsychro)](https://CRAN.R-project.org/package=ggpsychro)
<!-- badges: end -->

> 'ggplot2' extension for making psychrometric charts.

## Installation

<!-- You can install the released version of ggpsychro from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("ggpsychro") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("hongyuanjia/ggpsychro")
```

## Example

### The basic

Similar with ggplot2, the creation of a psychrometric chart using ggpsychro starts with function
`ggpsychro()`. You can specify the range of dry-bulb temperature and humidity
ratio using `tab_lim` and `hum_lim` and the unit system using `units`

```{r ggpsychro}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 50), units = "IP")
```

You can also create a mollier chart by setting `mollier` to `TRUE`.

```{r mollier}
ggpsychro(mollier = TRUE)
```

By default, only dry-bulb temperature, humidity ratio and the saturation line
are plotted.

The style of the saturation and area above it can be further styled using
`sat_style` and `mask_style`.

```{r style}
ggpsychro(mask_style = list(fill = "gray"), sat_style = list(color = "black", linetype = 2))
```

### Grid

ggpsychro introduces new ggplot geoms to add 5 more grids onto the base plot:

| Geom                 | Type                 |
| ---                  | ---                  |
| `geom_grid_relhum`   | Relative humidity    |
| `geom_grid_wetbulb`  | Wet-bulb temperature |
| `geom_grid_vappres`  | Vaper pressure       |
| `geom_grid_specvol`  | Specific volume      |
| `geom_grid_enthalpy` | Enthalpy             |

```{r grid, out.width = "100%", dpi = 300, fig.width = 15, fig.height = 10}
ggpsychro() +
    geom_grid_relhum() +
    geom_grid_wetbulb() +
    geom_grid_vappres() +
    geom_grid_specvol() +
    geom_grid_enthalpy()
```

Each of the new geom come along with a corresponding `scale_*` which can be used
to further customize the breaks of each grid

```{r grid-2, out.width = "100%", dpi = 300, fig.width = 15, fig.height = 10}
ggpsychro() +
    geom_grid_relhum(alpha = 1.0, label.alpha = 1.0, label.size = 6, label.fontface = 2) +
    scale_relhum(minor_breaks = NULL) +
    geom_grid_wetbulb(size = 1.0, color = "black", alpha = 1.0, label_loc = NA) +
    scale_wetbulb(breaks = seq(25, 30, by = 5), minor_breaks = NULL) +
    geom_grid_vappres(label.size = 5) +
    scale_vappres(breaks = seq(6000, 7000, by = 500), limits = c(6000, 7000)) +
    geom_grid_specvol() +
    scale_specvol(labels = NULL) +
    geom_grid_enthalpy()
```

### Stat

ggpsychro provides 5 new ggplot stats to use together with other common ggplot2
geoms::

| Stat            | Type                 |
| ---             | ---                  |
| `stat_relhum`   | Relative humidity    |
| `stat_wetbulb`  | Wet-bulb temperature |
| `stat_vappres`  | Vaper pressure       |
| `stat_specvol`  | Specific volume      |
| `stat_enthalpy` | Enthalpy             |

This makes it quick easy to add new elements.

Working together with ggplot2 orignal geoms is as simple as change `stat` to
your variable name of interest.

```{r stat-1}
library(eplusr) # for reading EPW data
epw <- read_epw(file.path(eplus_config(8.8)$dir, "WeatherData/USA_CO_Golden-NREL.724666_TMY3.epw"))

ggpsychro(epw$data()[month %in% 5:8]) +
    geom_grid_relhum() +
    geom_point(aes(dry_bulb_temperature, relhum = relative_humidity/100), stat = "relhum", alpha = 0.1) +
    facet_wrap(~month, labeller = as_labeller(function (x) paste0("Month: ", x)))
```

```{r stat-2}
ggpsychro(tdb_lim = c(10, 35), hum_lim = c(0, 30)) +
    geom_grid_relhum() +
    geom_grid_wetbulb() +
    # 18 wet-bulb line with dry-bulb from 20 - 30
    geom_line(aes(x = 20:30, wetbulb = 18), stat = "wetbulb")
```

## Author

Hongyuan Jia

## License

The project is released under the terms of MIT License.

Copyright © 2019-2020 Hongyuan Jia

## Contribute

Please note that the 'ggpsychro' project is released with a
[Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
