---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(ggpsychro)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    fig.align = "center",
    out.width = "60%",
    dev = "png",
    dpi = 100
)
```

# ggpsychro

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/hongyuanjia/ggpsychro/branch/master/graph/badge.svg)](https://codecov.io/gh/hongyuanjia/ggpsychro?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/ggpsychro)](https://CRAN.R-project.org/package=ggpsychro)
[![R-CMD-check](https://github.com/hongyuanjia/ggpsychro/workflows/R-CMD-check/badge.svg)](https://github.com/hongyuanjia/ggpsychro/actions)
<!-- badges: end -->

> 'ggplot2' extension for making psychrometric charts.

## Installation

<!-- You can install the released version of ggpsychro from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("ggpsychro") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("hongyuanjia/ggpsychro")
```

## Example

### The basic

Similar with ggplot2, the creation of a psychrometric chart using ggpsychro starts with function
`ggpsychro()`. You can specify the range of dry-bulb temperature and humidity
ratio using `tdb_lim` and `hum_lim` and the unit system using `units`.

```{r ggpsychro}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 50), units = "IP")
```

You can also create a mollier chart by setting `mollier` to `TRUE`.

```{r mollier}
ggpsychro(mollier = TRUE)
```

The coordinate system can also be added explicitly with `coord_psychro()`.

```{r coord}
ggpsychro() +
    coord_psychro(tdb_lim = c(10, 35), hum_lim = c(0, 30), units = "SI")
```

By default, ggpsychro draws dry-bulb temperature and humidity ratio grid lines,
the saturation curve, and the psychrometric reference grids.

The style of the saturation line and the mask area can be changed using
ggplot2 theme elements.

```{r style}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 50)) +
    theme(
        psychro.panel.grid.saturation = element_line(color = "black", linetype = 2),
        psychro.panel.mask = element_polygon(fill = "gray", color = NA)
    )
```

### Grid

ggpsychro renders psychrometric reference grids in `coord_psychro()`. The
default background grid is drawn without labels. The thin ggplot-style helpers
can make individual grids explicit and label their major lines:

| Geom                 | Type                 |
| ---                  | ---                  |
| `geom_grid_relhum`   | Relative humidity    |
| `geom_grid_wetbulb`  | Wet-bulb temperature |
| `geom_grid_vappres`  | Vapor pressure       |
| `geom_grid_specvol`  | Specific volume      |
| `geom_grid_enthalpy` | Enthalpy             |

```{r grid, out.width = "100%", dpi = 300, fig.width = 15, fig.height = 10}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 50)) +
    geom_grid_relhum() +
    geom_grid_wetbulb() +
    geom_grid_vappres() +
    geom_grid_specvol() +
    geom_grid_enthalpy()
```

Each grid comes with a corresponding `scale_*_continuous()` function for
customizing breaks and labels. The `geom_grid_*()` helpers do not create data
layers; they mark grids as visible or hidden and optionally override line and
label style.

```{r grid-2, out.width = "100%", dpi = 300, fig.width = 15, fig.height = 10}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 50)) +
    geom_grid_relhum(linewidth = 0.8, color = "black", label.size = 5) +
    scale_relhum_continuous(breaks = seq(25, 75, by = 25), minor_breaks = NULL) +
    geom_grid_wetbulb(color = "black", linewidth = 0.6, label = FALSE) +
    scale_wetbulb_continuous(breaks = seq(5, 30, by = 5), minor_breaks = NULL) +
    geom_grid_vappres(show = FALSE) +
    scale_specvol_continuous(labels = NULL) +
    geom_grid_enthalpy()
```

### Stat

ggpsychro provides 5 new ggplot stats to use together with other common ggplot2
geoms:

| Stat            | Type                 |
| ---             | ---                  |
| `stat_relhum`   | Relative humidity    |
| `stat_wetbulb`  | Wet-bulb temperature |
| `stat_vappres`  | Vapor pressure       |
| `stat_specvol`  | Specific volume      |
| `stat_enthalpy` | Enthalpy             |

This makes it quick and easy to add new elements.

Working together with ggplot2 original geoms is as simple as changing `stat` to
your variable name of interest.

```{r stat-1, message = FALSE, warning = FALSE}
library(eplusr) # for reading EPW data
epw <- read_epw(file.path(
    eplus_config(9.6)$dir,
    "WeatherData/USA_CO_Golden-NREL.724666_TMY3.epw"
))

ggpsychro(epw$data()[month %in% 5:8], tdb_lim = c(0, 50), hum_lim = c(0, 50)) +
    geom_grid_relhum() +
    geom_point(
        aes(dry_bulb_temperature, relhum = relative_humidity),
        stat = "relhum", color = "#0f766e", alpha = 0.24, size = 0.5, shape = 16
    ) +
    facet_wrap(~month, labeller = as_labeller(function(x) paste0("Month: ", x)))
```

```{r stat-2}
ggpsychro(tdb_lim = c(10, 35), hum_lim = c(0, 30)) +
    geom_grid_relhum() +
    geom_grid_wetbulb() +
    # 18 wet-bulb line with dry-bulb from 20 - 30
    geom_line(aes(x = 20:30, wetbulb = 18), stat = "wetbulb")
```

## Author

Hongyuan Jia

## License

The project is released under the terms of MIT License.

Copyright © 2019-2026 Hongyuan Jia

## Contribute

Please note that the 'ggpsychro' project is released with a
[Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
