---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(ggpsychro)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    fig.align = "center",
    out.width = "60%",
    dev = "png",
    dpi = 100
)
```

# ggpsychro

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/hongyuanjia/ggpsychro/branch/master/graph/badge.svg)](https://codecov.io/gh/hongyuanjia/ggpsychro?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/ggpsychro)](https://CRAN.R-project.org/package=ggpsychro)
[![R-CMD-check](https://github.com/hongyuanjia/ggpsychro/workflows/R-CMD-check/badge.svg)](https://github.com/hongyuanjia/ggpsychro/actions)
<!-- badges: end -->

> 'ggplot2' extension for making psychrometric charts.

## Installation

<!-- You can install the released version of ggpsychro from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("ggpsychro") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("hongyuanjia/ggpsychro")
```

## Example

The creation of a psychrometric chart starts with `ggpsychro()`. The result is a
ggplot object, so grids, stats, geoms, scales, and themes can be added with the
same `+` workflow used by ggplot2.

```{r ggpsychro, fig.alt = "Minimal psychrometric chart with dry-bulb temperature, humidity ratio, and reference grids."}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 30)) +
    psychro_preset("minimal")
```

State points and process lines can be supplied with dry-bulb temperature plus one
psychrometric property, such as relative humidity.

```{r psychro-process, out.width = "75%", fig.alt = "Psychrometric process line with four state points on a minimal chart."}
process <- data.frame(
    dry_bulb_temperature = c(18, 23, 28, 31),
    relative_humidity = c(70, 55, 45, 55)
)

ggpsychro(tdb_lim = c(0, 40), hum_lim = c(0, 25)) +
    psychro_preset("minimal") +
    geom_psychro_process(
        aes(tdb = dry_bulb_temperature, relhum = relative_humidity),
        data = process,
        colour = "#0f766e", linewidth = 1,
        arrow = grid::arrow(length = grid::unit(0.08, "inches"))
    ) +
    stat_psychro_state(
        aes(tdb = dry_bulb_temperature, relhum = relative_humidity),
        data = process,
        colour = "#0f766e", size = 2
    )
```

## Learn more

The longer examples live on the pkgdown site:

* [Get started](https://hongyuanjia.github.io/ggpsychro/articles/ggpsychro.html)
* [Chart grids and styles](https://hongyuanjia.github.io/ggpsychro/articles/grids-and-styles.html)
* [Plotting psychrometric data](https://hongyuanjia.github.io/ggpsychro/articles/plotting-data.html)
* [Zones and processes](https://hongyuanjia.github.io/ggpsychro/articles/zones-and-processes.html)

## Author

Hongyuan Jia

## License

The project is released under the terms of MIT License.

Copyright © 2019-2026 Hongyuan Jia

## Contribute

Please note that the 'ggpsychro' project is released with a
[Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
