---
title: "Comfort overlays"
---

```{r setup, include = FALSE}
library(ggpsychro)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.align = "center",
    out.width = "100%",
    fig.width = 9,
    fig.height = 6,
    dev = "png",
    dpi = 120
)
```

Comfort overlay layers evaluate thermal comfort models on the current
psychrometric panel. They inherit the chart units, pressure, limits, and
Mollier orientation from `ggpsychro()`, so comfort regions can be composed with
the same `+` workflow as other ggplot layers.

Use the layer that matches the question you want the chart to answer:

| Question | Layer |
|---|---|
| How does PMV vary across the chart? | `geom_comfort_pmv()` |
| Where are PMV-based ASHRAE 55 or EN 15251 comfort zones? | `geom_comfort_pmv(standard = ...)` |
| How does SET vary across the chart? | `geom_comfort_set()` |
| What adaptive comfort range applies for a running outdoor temperature? | `geom_comfort_adaptive()` |
| Which Outdoor Work Heat Index category applies? | `geom_comfort_heat_index()` |
| Which Givoni-Milne strategy regions apply? | `geom_comfort_givoni()` |
| What comfort metric applies to measured state points? | `stat_comfort_state()` |

## PMV overlays

The default comfort model is ISO 7730 PMV. `geom_comfort_pmv()` draws filled PMV
bands, labelled PMV contours, and optional PMV-based standard zones.
`scale_fill_comfort_pmv()` applies a PMV-centered diverging palette.

```{r pmv-overlay, fig.alt = "Psychrometric chart with filled PMV bands and labelled PMV contour lines."}
ggpsychro(tdb_lim = c(5, 40), hum_lim = c(0, 24)) +
    psychro_preset("minimal") +
    geom_comfort_pmv(
        contour_levels = seq(-3, 3, by = 0.5),
        n = c(70, 48)
    ) +
    scale_fill_comfort_pmv(name = "PMV")
```

### Resolution and rendering modes

Comfort layers use `n` as the rendering-resolution control. For filled bands,
pass one value for both dry-bulb and humidity-ratio directions or two values as
`c(tdb, humratio)`. PMV contour curves use the first value because they trace
constant-PMV roots along one sampling direction. Lower values build faster and
are useful during exploration; higher values give smoother final graphics.

Use `band_render = "band"` for filled regions and `band_render = "tile"` for
direct sampled cells. For PMV bands, `band_method = "auto"` uses root-traced
boundaries; `band_method = "isoband"` uses the cheaper gridded approximation.

Default resolutions are chosen per model so the common examples build quickly
while keeping smooth visible boundaries:

| Layer/model | Default `n` | Notes |
|---|---:|---|
| `geom_comfort_pmv()` bands | `c(360, 220)` | PMV contours and standard boundaries use the first value. |
| `geom_comfort_set()` | `c(80, 50)` | SET bands and contours use a sampled grid. |
| `geom_comfort_adaptive()` | `c(240, 160)` | Adaptive zones are sampled and clipped to the chart. |
| `geom_comfort_heat_index()` | `c(160, 100)` | Heat-index categories share a cached sampled grid. |

## Standard comfort zones

For PMV-based standards, pass a standard helper to `geom_comfort_pmv()`. The
standard layer draws the filled zone, PMV boundary contours, and labels.

```{r ashrae55-zone, fig.alt = "Psychrometric chart with the PMV-based ASHRAE 55 2017 comfort zone."}
ggpsychro(tdb_lim = c(5, 35), hum_lim = c(0, 24)) +
    psychro_preset("minimal") +
    geom_comfort_pmv(
        standard = comfort_pmv_ashrae55(),
        bands = FALSE,
        contours = FALSE,
        n = 140
    )
```

```{r en15251-zone, fig.alt = "Psychrometric chart with PMV-based EN 15251 2007 comfort bands."}
ggpsychro(tdb_lim = c(5, 35), hum_lim = c(0, 24)) +
    psychro_preset("minimal") +
    geom_comfort_pmv(
        standard = comfort_pmv_en15251(),
        bands = FALSE,
        contours = FALSE,
        n = 140
    )
```

## SET and adaptive models

Create reusable model objects with `comfort_model_*()` helpers. SET overlays
default to the `"set"` metric, while adaptive models default to
`"acceptability"`. For continuous metrics such as SET, use filled bands with
labelled contours so the field remains visible without losing the level
boundaries. `geom_comfort_set(contours = TRUE, labels = TRUE)` draws contour
labels inline and leaves a gap in each labelled contour where the label is
placed.

```{r set-overlay, fig.alt = "Psychrometric chart with filled SET bands and labelled SET contour lines."}
set_model <- comfort_model_set()

ggpsychro(tdb_lim = c(15, 30), hum_lim = c(0, 20)) +
    psychro_preset("minimal") +
    geom_comfort_set(
        model = set_model,
        band_levels = seq(14, 32, by = 2),
        contours = TRUE,
        contour_levels = seq(22, 30, by = 2),
        labels = TRUE,
        n = c(70, 42),
        alpha = 0.32
    ) +
    scale_fill_viridis_c("SET (deg C)", option = "C")
```

```{r adaptive-zone, fig.alt = "Psychrometric chart with an adaptive comfort acceptability zone."}
ggpsychro(tdb_lim = c(15, 30), hum_lim = c(0, 20)) +
    psychro_preset("minimal") +
    geom_comfort_adaptive(
        t_running = 20,
        fill = "#22c55e",
        alpha = 0.18,
        colour = "#166534",
        linewidth = 0.7
    )
```

Adaptive comfort zones are temperature bands rather than humidity-dependent
fields, so a filled adaptive zone is expected to look like a vertical band
instead of a curved humidity field. In the example above, `t_running = 20`
gives an ASHRAE 55 80% acceptability interval from 20.5 to 27.5 degrees C;
because `tr` defaults to `tdb`, the zone is clipped only by the current chart
limits and saturation boundary.

## Outdoor Heat Index

Heat Index overlays use `comfort_model_heat_index()` and
`geom_comfort_heat_index()` to draw Outdoor Work Heat Index categories. The
model uses a NOAA-style Heat Index calculation internally in Fahrenheit, while
the layer follows the parent chart's unit system.

```{r heat-index-overlay, fig.alt = "Psychrometric chart with Outdoor Work Heat Index caution and danger categories."}
ggpsychro(tdb_lim = c(20, 45), hum_lim = c(0, 35)) +
    psychro_preset("minimal") +
    geom_comfort_heat_index(n = c(70, 42), alpha = 0.5)
```

## Givoni-Milne strategy overlay

Givoni-Milne overlays are strategy zones rather than a single continuous
comfort metric. Create a strategy with `comfort_strategy_givoni()` and draw it
with `geom_comfort_givoni()`. The default adaptive variant shifts the comfort
anchor from mean outdoor temperature and shows that temperature as a dashed
marker. Use `variant = "fixed"` for the commonly cited 1979 comfort anchor of
20 to 25.5 degrees C and 20% to 80% relative humidity, with the hot-humid corner
clipped. For project- or climate-specific assumptions, set `tdb_range` and
`relhum_range` on the fixed variant to use those ranges as the drawing anchor.
Treat these regions as a pre-design screening aid: high-mass and night
ventilation regions need daily profiles, nighttime conditions, and building
assumptions before they can be used to count comfort hours.

```{r givoni-zones, fig.alt = "Psychrometric chart with Givoni-Milne strategy zones and a mean outdoor temperature marker."}
givoni <- comfort_strategy_givoni(mean_outdoor = 22)

ggpsychro(tdb_lim = c(-10, 50), hum_lim = c(0, 35)) +
    psychro_preset("minimal") +
    geom_comfort_givoni(givoni, alpha = 0.45)
```

```{r givoni-fixed, fig.alt = "Psychrometric chart with fixed Givoni-Milne strategy zones."}
fixed_givoni <- comfort_strategy_givoni(variant = "fixed", mean_outdoor = NULL)

ggpsychro(tdb_lim = c(-10, 50), hum_lim = c(0, 35)) +
    psychro_preset("minimal") +
    geom_comfort_givoni(fixed_givoni, alpha = 0.45)
```

```{r givoni-custom-anchor, fig.alt = "Psychrometric chart with custom Givoni-Milne comfort anchor ranges."}
custom_givoni <- comfort_strategy_givoni(
    variant = "fixed",
    mean_outdoor = NULL,
    tdb_range = c(22, 27),
    relhum_range = c(30, 70)
)

ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 35)) +
    psychro_preset("minimal") +
    geom_comfort_givoni(custom_givoni, alpha = 0.45)
```

Use `zone_style` with `element_givoni_zone()` to highlight individual strategy
zones without changing the strategy geometry.

```{r givoni-zone-style, fig.alt = "Psychrometric chart with styled Givoni comfort, natural ventilation, winter, and air-conditioning zones."}
ggpsychro(tdb_lim = c(-10, 50), hum_lim = c(0, 35)) +
    psychro_preset("minimal") +
    geom_comfort_givoni(
        givoni,
        zone_style = list(
            comfort = element_givoni_zone(
                fill = "#66D27A", colour = "#1F5F2D", alpha = 0.35
            ),
            natural_ventilation = element_givoni_zone(
                fill = "#B6E3FF", colour = "#2F6FB0", alpha = 0.25
            ),
            winter = element_givoni_zone(
                colour = "#A14D00", linetype = "dashed"
            ),
            air_conditioning = element_givoni_zone(
                colour = "#8B1E3F", linewidth = 1.2
            )
        )
    )
```

## State metrics

`stat_comfort_state()` evaluates the model at supplied psychrometric states.
The computed comfort fields can be used with `after_stat()`.

```{r state-metrics, fig.alt = "Psychrometric chart with measured state points coloured by computed PMV."}
states <- data.frame(
    tdb = c(22, 24, 26, 28),
    relhum = c(45, 50, 55, 60)
)

ggpsychro(states, tdb_lim = c(15, 32), hum_lim = c(0, 22)) +
    psychro_preset("minimal") +
    stat_comfort_state(
        aes(tdb = tdb, relhum = relhum, colour = after_stat(pmv)),
        size = 3
    ) +
    scale_colour_gradient2(
        "PMV",
        low = "#3B5FFF",
        mid = "#4A4A4A",
        high = "#FF3B30",
        midpoint = 0
    )
```

## Mollier and IP charts

Comfort layers follow the parent chart's coordinate system. Use
`mollier = TRUE` for Mollier orientation or `units = "IP"` for IP inputs.

```{r mollier-comfort-overlay, fig.alt = "Mollier psychrometric chart with a PMV comfort overlay and labelled PMV lines."}
ggpsychro(tdb_lim = c(15, 30), hum_lim = c(0, 20), mollier = TRUE) +
    psychro_preset("minimal") +
    geom_comfort_pmv(
        contour_levels = seq(-2, 2, by = 1),
        n = c(50, 30)
    ) +
    scale_fill_comfort_pmv(name = "PMV")
```

Use manual psychrometric zones when the desired region is a project-specific
operating envelope rather than a thermal comfort model. See [Zones and
processes](zones-and-processes.html) for those layers.
