---
title: "Comfort overlays"
---

```{r setup, include = FALSE}
library(ggpsychro)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.align = "center",
    out.width = "100%",
    fig.width = 9,
    fig.height = 6,
    dev = "png",
    dpi = 120
)
```

Comfort overlay layers evaluate thermal comfort models on the current
psychrometric panel. They inherit the chart units, pressure, limits, and
Mollier orientation from `ggpsychro()`, so comfort regions can be composed with
the same `+` workflow as other ggplot layers.

## PMV overlays

The default comfort model is ISO 7730 PMV. `geom_comfort_overlay()` draws filled
PMV bands, `scale_fill_comfort_pmv()` applies a PMV-centered diverging palette,
and `geom_comfort_pmv_lines()` adds labelled PMV curves.

```{r pmv-overlay, fig.alt = "Psychrometric chart with filled PMV bands and labelled PMV curve lines."}
ggpsychro(tdb_lim = c(5, 40), hum_lim = c(0, 24)) +
    psychro_preset("minimal") +
    geom_comfort_overlay(n = c(70, 48), gap = 0) +
    scale_fill_comfort_pmv(name = "PMV") +
    geom_comfort_pmv_lines(levels = seq(-3, 3, by = 0.5), n = 140)
```

## Standard comfort zones

For PMV-based standards, `geom_comfort_standard_zone()` draws the filled zone,
PMV boundary curves, and labels from a standard helper.

```{r ashrae55-zone, fig.alt = "Psychrometric chart with the PMV-based ASHRAE 55 2017 comfort zone."}
ggpsychro(tdb_lim = c(5, 35), hum_lim = c(0, 24)) +
    psychro_preset("minimal") +
    geom_comfort_standard_zone(comfort_standard_ashrae55_2017(), n = 140)
```

```{r en15251-zone, fig.alt = "Psychrometric chart with PMV-based EN 15251 2007 comfort bands."}
ggpsychro(tdb_lim = c(5, 35), hum_lim = c(0, 24)) +
    psychro_preset("minimal") +
    geom_comfort_standard_zone(comfort_standard_en15251_2007(), n = 140)
```

## SET and adaptive models

Create reusable model objects with `comfort_model_*()` helpers. SET overlays
default to the `"set"` metric, while adaptive models default to
`"acceptability"`. For continuous metrics such as SET, use filled bands with
contours so the field remains visible without losing the level boundaries.

```{r set-overlay, fig.alt = "Psychrometric chart with filled SET bands and SET contour lines."}
set_model <- comfort_model_set()

ggpsychro(tdb_lim = c(15, 30), hum_lim = c(0, 20)) +
    psychro_preset("minimal") +
    geom_comfort_overlay(
        model = set_model,
        metric = "set",
        levels = seq(14, 32, by = 2),
        n = c(70, 42),
        alpha = 0.32
    ) +
    scale_fill_viridis_c("SET (deg C)", option = "C") +
    geom_comfort_contour(
        model = set_model,
        metric = "set",
        breaks = seq(18, 30, by = 2),
        n = c(70, 42),
        colour = "#2F2F2F",
        linewidth = 0.7
    )
```

```{r adaptive-zone, fig.alt = "Psychrometric chart with an adaptive comfort acceptability zone."}
adaptive_model <- comfort_model_adaptive(t_running = 20)

ggpsychro(tdb_lim = c(15, 30), hum_lim = c(0, 20)) +
    psychro_preset("minimal") +
    geom_comfort_zone(
        model = adaptive_model,
        fill = "#22c55e",
        alpha = 0.18,
        colour = "#166534",
        linewidth = 0.7
    )
```

Adaptive comfort zones are temperature bands rather than humidity-dependent
fields. In the example above, `t_running = 20` gives an ASHRAE 55 80%
acceptability interval from 20.5 to 27.5 degrees C; because `tr` defaults to
`tdb`, the zone appears as a vertical band clipped to the current chart limits.

## State metrics

`stat_comfort_state()` evaluates the model at supplied psychrometric states.
The computed comfort fields can be used with `after_stat()`.

```{r state-metrics, fig.alt = "Psychrometric chart with measured state points coloured by computed PMV."}
states <- data.frame(
    tdb = c(22, 24, 26, 28),
    relhum = c(45, 50, 55, 60)
)

ggpsychro(states, tdb_lim = c(15, 32), hum_lim = c(0, 22)) +
    psychro_preset("minimal") +
    stat_comfort_state(
        aes(tdb = tdb, relhum = relhum, colour = after_stat(pmv)),
        size = 3
    ) +
    scale_colour_gradient2(
        "PMV",
        low = "#3B5FFF",
        mid = "#4A4A4A",
        high = "#FF3B30",
        midpoint = 0
    )
```

## Mollier and IP charts

Comfort layers follow the parent chart's coordinate system. Use
`mollier = TRUE` for Mollier orientation or `units = "IP"` for IP inputs.

```{r mollier-comfort-overlay, fig.alt = "Mollier psychrometric chart with a PMV comfort overlay and labelled PMV lines."}
ggpsychro(tdb_lim = c(15, 30), hum_lim = c(0, 20), mollier = TRUE) +
    psychro_preset("minimal") +
    geom_comfort_overlay(n = c(50, 30)) +
    scale_fill_comfort_pmv(name = "PMV") +
    geom_comfort_pmv_lines(levels = seq(-2, 2, by = 1), n = 100)
```

Use manual psychrometric zones when the desired region is a project-specific
operating envelope rather than a thermal comfort model. See [Zones and
processes](zones-and-processes.html) for those layers.
