---
title: "Get started"
---

```{r setup, include = FALSE}
library(ggpsychro)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.align = "center",
    out.width = "100%",
    fig.width = 8,
    fig.height = 5,
    dev = "png",
    dpi = 120
)
```

ggpsychro creates psychrometric charts with the same additive grammar used by
ggplot2. Start with `ggpsychro()`, choose the chart limits and units, and then
add grids, scales, stats, geoms, and themes with `+`.

```{r basic-chart, fig.alt = "Default psychrometric chart with dry-bulb temperature and humidity ratio axes."}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 30))
```

## Chart limits and units

`tdb_lim` controls the dry-bulb temperature range. `hum_lim` controls the
displayed humidity ratio range. SI units are used by default.

```{r si-chart, fig.alt = "SI psychrometric chart limited to 10 to 35 degrees C and 0 to 25 g per kg humidity ratio."}
ggpsychro(tdb_lim = c(10, 35), hum_lim = c(0, 25), units = "SI")
```

Use `units = "IP"` for IP charts. The dry-bulb limits are in degree F and the
humidity-ratio limits are in grains per pound of dry air.

```{r ip-chart, fig.alt = "IP psychrometric chart limited to 32 to 122 degrees F and 0 to 220 grains per pound humidity ratio."}
ggpsychro(tdb_lim = c(32, 122), hum_lim = c(0, 220), units = "IP")
```

## Mollier charts

Set `mollier = TRUE` to swap the chart into Mollier orientation.

```{r mollier-chart, fig.alt = "Mollier chart with humidity ratio on the x axis and dry-bulb temperature on the y axis."}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 30), mollier = TRUE)
```

## Explicit coordinates

`ggpsychro()` sets up the psychrometric coordinate system for you. If you are
building up a plot in smaller pieces, `coord_psychro()` can also be added
explicitly.

```{r coord-psychro, fig.alt = "Psychrometric chart created by adding coord_psychro explicitly."}
ggpsychro() +
    coord_psychro(tdb_lim = c(10, 35), hum_lim = c(0, 25), units = "SI")
```

## Add ggplot layers

Because the result is a ggplot object, regular ggplot layers work as usual.

```{r add-points, fig.alt = "Minimal psychrometric chart with four state points added as a scatter layer."}
states <- data.frame(
    dry_bulb_temperature = c(18, 23, 28, 31),
    humidity_ratio = c(8, 9.5, 10.5, 13)
)

ggpsychro(states, aes(dry_bulb_temperature, humidity_ratio),
    tdb_lim = c(0, 40), hum_lim = c(0, 25)
) +
    psychro_preset("minimal") +
    geom_point(color = "#0f766e", size = 2)
```

For data recorded as relative humidity, wet-bulb temperature, vapor pressure,
specific volume, or enthalpy, see [Plotting psychrometric
data](plotting-data.html).
