---
title: "Chart grids and styles"
---

```{r setup, include = FALSE}
library(ggpsychro)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.align = "center",
    out.width = "100%",
    fig.width = 9,
    fig.height = 6,
    dev = "png",
    dpi = 120
)
```

ggpsychro draws the dry-bulb grid, humidity-ratio grid, saturation curve, and
psychrometric reference grids through `coord_psychro()`. The default reference
grids are drawn without labels. Add `geom_grid_*()` helpers when a grid should
be explicit and labelled.

## Presets

`psychro_preset()` applies a named collection of scales, labelled reference
grids, and a matching theme. The built-in presets are `"ashrae"` and
`"minimal"`.

```{r ashrae-preset, fig.width = 12, fig.height = 8, fig.alt = "ASHRAE-style psychrometric chart with labelled reference grids."}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 30)) +
    psychro_preset("ashrae")
```

```{r minimal-preset, fig.width = 12, fig.height = 8, fig.alt = "Minimal psychrometric chart with selected labelled reference grids."}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 40)) +
    psychro_preset("minimal")
```

Use `labels = FALSE` when you want the preset's grid visibility and theme but no
grid labels.

```{r preset-no-labels, fig.alt = "ASHRAE-style psychrometric chart with reference grid labels hidden."}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 30)) +
    psychro_preset("ashrae", labels = FALSE)
```

## Reference grid helpers

The reference grid helpers are:

| Helper               | Property             |
|----------------------|----------------------|
| `geom_grid_relhum()` | Relative humidity    |
| `geom_grid_wetbulb()` | Wet-bulb temperature |
| `geom_grid_vappres()` | Vapor pressure       |
| `geom_grid_specvol()` | Specific volume      |
| `geom_grid_enthalpy()` | Enthalpy            |

```{r all-grids, fig.width = 12, fig.height = 8, fig.alt = "Psychrometric chart with relative humidity, wet-bulb, vapor pressure, specific volume, and enthalpy grids."}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 40)) +
    geom_grid_relhum() +
    geom_grid_wetbulb() +
    geom_grid_vappres() +
    geom_grid_specvol() +
    geom_grid_enthalpy()
```

These helpers do not add data layers. They update the psychrometric coordinate
metadata so the panel background can draw or hide the selected grid.

## Breaks, labels, and style

Each reference grid has a matching `scale_*_continuous()` function. Use these
scales to customize breaks, minor breaks, limits, and labels.

```{r custom-grid, fig.width = 12, fig.height = 8, fig.alt = "Psychrometric chart with custom grid breaks, labels, and line styles."}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 40)) +
    geom_grid_relhum(linewidth = 0.8, color = "black", label.size = 5) +
    scale_relhum_continuous(breaks = seq(25, 75, by = 25), minor_breaks = NULL) +
    geom_grid_wetbulb(color = "black", linewidth = 0.6, label = FALSE) +
    scale_wetbulb_continuous(breaks = seq(5, 30, by = 5), minor_breaks = NULL) +
    geom_grid_vappres(show = FALSE) +
    scale_specvol_continuous(labels = NULL) +
    geom_grid_enthalpy()
```

Line style arguments go directly to `ggplot2::element_line()`. Label style
arguments use a `label.` prefix, such as `label.size`, `label.color`, and
`label.vjust`.

## Theme elements

ggpsychro adds psychrometric theme elements for panel masks and reference
grids. Use these with regular `theme()` calls.

```{r theme-elements, fig.alt = "Psychrometric chart with custom saturation line and gray panel mask theme elements."}
ggpsychro(tdb_lim = c(0, 50), hum_lim = c(0, 30)) +
    geom_grid_relhum(label = FALSE) +
    theme(
        psychro.panel.grid.saturation = element_line(
            color = "black", linetype = 2
        ),
        psychro.panel.mask = element_polygon(fill = "gray90", color = NA)
    )
```
